Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2019-09-03 hpmoon-asoco.tex
Adding tag and final version submitted

\newenvironment{bluecolor}{\par\color{blue}}{\par}

@@

%%%NEW AND SHORTED
Parallel island-model co-evolutionary algorithms are well-known
methods, suitable for dealing with large multi-objective optimization
problems. \textcolor{blue}{This paper proposes a version of these algorithms where} each island modifies a fragment of the
chromosome that encodes a possible solution to the
problem. \textcolor{blue}{The objective of this paper is} to
demonstrate that automatically setting the size of the overlapping
fragments depending on the number of islands obtains better results
than using \textcolor{blue}{a fixed overlapping} size. \textcolor{blue}{This method has been
  compared} to other parallel evolutionary techniques considering a
different number of islands, chromosome sizes and benchmarks. The
analysis of the obtained experimental results, by using different
metrics, shows that our approach can provide statistically significant
improvements with respect to the base algorithm in high-dimensional,
un-decomposable, multi-objective problems. This opens a very promising
line to automatically adapt the overlapping sizes in this kind of
algorithms.

@@

% Maybe add a sentence explaining what problems are decomposable and which ones are not. - JJ
% Pablo: ok, done in blue
\begin{bluecolor}
One of the methods that are most extensively used  to deal with
high-dimensional, multi-objective optimization problems (MOPs) are
distributed co-evolutionary algorithms \cite{Gong15models}. However,
only decomposable problems \textcolor{blue}{--that is, those whose  domain can be divided into subsets to calculate their partial solutions and combine them to form the solution of the original problem--,} could strictly be implemented in parallel
this way. A new parallel island-model evolutionary algorithm approach
\cite{Alba02Parallelism} is used here to deal with un-decomposable
problems: \textcolor{blue}{instead of allocating disjoint subsets of components to the islands hosted by the available processors, it applies the variation operators only to the components  of the subset allocated to a given population, while keeping the rest of the components of the whole individual unaltered}. This avoids extra steps (and
subsequent communication overhead) to recompose the complete
individuals from other islands before evaluation. 
\end{bluecolor}

%1. MOP ARE
In MOPs, several \textcolor{blue}{independent}
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives. The objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as possible. In order to do that\textcolor{blue}{,} the
exploration of the search space must be very efficient, trying to
cover \textcolor{blue}{all} possible exploration paths at the same time.

%2 COMPUTATION -> PARALLEL
This aspect of searching for non-dominated solutions, as well as, in
many cases, the size of the search space itself, implies a high demand
of computational time, leading to the proposal of parallel and distributed methods to solve them
\citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268},
%3 ONE WAY IS TO USE MOEAS BECAUSE
one of which \textcolor{blue}{is} evolutionary algorithms (EAs)
\citep{DBLP:series/ncs/EibenS15}. EAs are bio-inspired meta-heuristics
that can be effectively used to find nearly optimal solutions for
optimization problems and are easily parallelizable, since they are
population-based. This means that creating a functionally equivalent parallel
version is as easy as distributing the population in communicating
computing nodes, which are usually called islands. The fact that
they are population-based helps find many solutions that compose the
Pareto front simultaneously.

Usually\textcolor{blue}{,} an EA starts by generating \textcolor{blue}{a \emph{population}, that is, a set of random solutions}, following a user-defined description. Then, it
evaluates each candidate solution, called \emph{individual}, assigning
it a \emph{fitness} \textcolor{blue}{vector, in the case of MOPs} that describes how good the individual is
\textcolor{blue}{solving} the target problem. New solutions are then generated by
the application of \emph{operators} that either mutate a single
solution or recombine different existing solutions. After each
iteration, called \emph{generation}, the least fit individuals are
removed, and the process continues until a user-defined stop condition
is met.

EAs \textcolor{blue}{were} extended to multi-objective optimization through quite a number of Multi-Objective Evolutionary Algorithms (MOEAs) \cite{TalbiUnified2018}.
One of the best known and \textcolor{blue}{highly} referenced MOEAs
\citep{Dorronsoro13superlinear} is the {\em 
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, which are divided into different ranks, based on 
dominance, to be selected for the next generations.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel,Gong15models}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies \textcolor{blue}{using} different distribution
and sharing mechanisms, as there exists a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}. \textcolor{blue}{This is also the case with the NSGA-II.} % Is that the case with NSGA? - JJ
% Please check the last question - JJ Pablo: done

Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}, 
meaning that the
MOEAs need to deal with large individuals \textcolor{blue}{representing those variables} and spend significant extra
time \textcolor{blue}{applying the} crossover, mutation and migration \textcolor{blue}{operators}.
% If this is the only justification, it should have been assessed at the end. It would be better if it was qualified, in the sense that more exploration is needed or something like that - JJ
Different authors have 
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. % Why should it improve quality? - JJ Pablo: look at the 2 new paragraphs sent by Julio below (between \begin{bluecolor})
In this aspect, the
co-evolution model is a dimension-distributed model where a 
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. \textcolor{blue}{To evolve the individuals of a subpopulation, the processor where it has been allocated can operate according to two main alternatives.} One
example of the application of the \textcolor{blue}{first alternative} was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs \textcolor{blue}{different alternatives for the recombination/cooperation of the subpopulations returned by worker processes}. A high
dimensional problem was used to compare these 
alternatives\textcolor{blue}{; also, i}n \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although  both papers \textcolor{blue}{described}
significant speedups, only a \textcolor{blue}{small} number of nodes were
employed (8). However, when the number of islands is increased, the division
of \textcolor{blue}{the portion of chromosome assigned to a given island} becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time. \textcolor{blue}{Moreover, in this first alternative, each processor only works with the variables allocated to it of the individuals in its population. Thus, only whenever the cost function is decomposable in such a way that the variables allocated to different processors are not interdependent, the evaluation of the fitness of an individual in the population can be done by composing the cost evaluated separately by each processor from the values of the variables allocated to it. Otherwise, it has to be determined after composing the whole solution from the components allocated to each processor. In fact, different strategies to define the way to distribute the components into groups to be allocated to the available processors are required. They have been mainly defined according to three categories \cite{Jia19Coevo}: (1) static partitions of variables into groups; (2) random dynamic grouping that changes the variables among groups along the optimization process; and (3) learning-based dynamic grouping to distribute the components by applying techniques to detect the interdependence among variables along the process. }

@@

Another approach used to solve this kind of problems is the {\em
  SeLective operator Application} (SLA), presented in this paper, \textcolor{blue}{can be considered as the second alternative to evolve the individuals of a subpopulation}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows \textcolor{blue}{dealing} with un-decomposable problems.

@@

In our previous work \citep{Garcia16hpmoon} \textcolor{blue}{this method was used} in a
preliminary experimental setup, although it is worth mentioning \textcolor{blue}{that its name and acronym is established} in this paper for the first time.
\textcolor{blue}{In that paper, it was shown}
that applying the variation operators only \textcolor{blue}{to} specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, \textcolor{blue}{the results showed} that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. 
This
motivate\textcolor{blue}{d} us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more \textcolor{blue}{extensive parameter evaluation}. %Again, "testing more stuff" is not a good motivation. You should try to address a specific research question that might have been raised in the previous paper - JJ
Besides comparing previous methods in this new experimental setup,  
a new method that automatically sets the size of the overlapped
sections depending on the number of available islands \textcolor{blue}{is proposed in this paper, and compared} with previous versions.

@@

Results also show that the relation between the number of islands
(subpopulations) and the length of the chromosome (number of decision
variables) is also a relevant factor to determine the most efficient
alternative to \textcolor{blue}{distribute} the decision variables: the higher the
dimensionality and the overlapping, the greater the benefit of
distribution in a large number of islands. This opens up a new
challenge to be addressed in the future, by experimenting with new
ways to calculate the overlapping sizes depending on different factors
and parameters.

The rest of the paper is organized as follows: after the State of the Art in distributed and co-evolutionary MOEAs, 
the \textcolor{blue}{proposed} methodology and the compared algorithms are described in Section \ref{sec:methodology}. 
Then, the results of the experiments are presented (Section \ref{sec:results}), Finally, the conclusions and future work are discussed.

@@

%FERGU: Texto que me ha pasado Julio
%\textcolor{blue}{Multi-objective optimization evolutionary algorithms (MOEAs) can take advantage of parallel and distributed processing \cite{Gong15models} not only to improve the quality of the solutions found, but also to speed up the execution times. Among the functional and data decomposition alternatives to parallelize an algorithm, parallel evolutionary algorithms usually implement data decomposition by identifying tasks that have to be repeatedly executed on different data. For example, in an evolutionary algorithm, the evaluation of the fitness function and the application of mutation operators to the individuals of the population can be independently applied to each individual. Moreover, in the context of data decomposition, two main alternatives can be distinguished. In the first one, the individuals of the population are distributed among several processes that independently compute their corresponding fitness functions. Once each process completes the fitness evaluation of the individuals allocated to it, another process applies the corresponding evolutionary operators to the individuals of the whole population to determine the population for the next generation. Then, the population is distributed again among the processes to obtain the new values of the fitness and so on until the end condition (a given number of generations, or a solution quality level) is met. It is clear that this alternative does not modify the convergence characteristics of the corresponding sequential evolutionary algorithm, although to complete each iteration communication is required between the process that applies the evolutionary operators to the population and the processes that evaluate the fitness values of the population. In the second alternative, several parallel processes execute iterations of the corresponding evolutionary algorithm on different subpopulations. In this case, the convergence behaviour could be different in sequential and parallel versions of the algorithms as the different subpopulations evolve in parallel and only exchange information about their individuals after completing a given number of generations. The number of individuals in the subpopulations (the grain size of the parallel procedure), and the characteristics of the communication (frequency of communication, character synchronous or asynchronous of the communication, characteristics of the communicated data, etc.) determine the performance of the parallel procedure. This second parallelization alternative (concurrent execution of evolutionary algorithms over multiple subpopulations) is the one considered in this paper. It can be implemented either by a master-worker approximation or by assigning each subpopulation to a different processor and communicating this processor with the others when it is required. In this second approach, according to the size of the grain (number of individuals in the subpopulations), two main alternatives can be also distinguished: the island parallel approach (coarse grain) and the cellular or diffusion approach (fine grain). In the master-worker approximation, the master distributes the subpopulations among the workers that communicate through the master. This way, as the communications increase (with the number of subpopulations and/or the frequency of communications among subpopulations) the master could become a bottleneck that reduces the efficiency of the parallel procedure.}

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other \textcolor{blue}{kinds} of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, the authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population into elite and search sub-populations \citep{Wang09parallel}, or separating it \textcolor{blue}{into} processors by objective \citep{Xiao03specialized}. For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector. Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Many real-world problems \textcolor{blue}{require} high-dimensional data \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality. Other authors, such as Cao et al. \cite{CaoZLL17} separate the decision variables into several subpopulations after performing an interdependence analysis, but using a fixed number of processors.

The approach described by Dorronsoro et al. has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but implementing a
master-slave method that splits the population into several processors. As in previous works, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper \textcolor{blue}{up to 128 islands have been used}.

Giving the responsibility of different parts of the search space to
different modules, has also been explored lately by some
researchers. For example, some authors focus on the solution space
\textcolor{blue}{modelling} the PF by dividing it into different clusters, allowing an
easily parallelizable {\em divide and conquer} approach
\citep{cheng2015adaptive}. 
Even other types of metaheuristics, 
such as the Multi-Objective Ant Colony Optimization (MOACO) \cite{Mora13paretobased}
or Artificial Bee Colony Algorithm \cite{LUO2017235}, 
can rely on island models to divide the search space depending on 
the node to achieve better results.

In our previous work \citep{Garcia16hpmoon} some of
the aforementioned ideas \textcolor{blue}{were used} to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections,  \textcolor{blue}{with} $P$ being the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time, both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. \textcolor{blue}{It was} discovered \textcolor{blue}{in that paper} that the
performance using one or another method depends on \textcolor{blue}{both} the number of
sections \textcolor{blue}{individual's section used by each island} and the number of islands. This
motivated us to find a new automatic method to select this number of
sections of the chromosome to \textcolor{blue}{be used}, depending on the number of
islands. Besides, previous experiments were performed on a single
processor \textcolor{blue}{with a} synchronous island model and a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and \textcolor{blue}{this time}, the experiments have been run on a parallel
cluster. 
Therefore, at the same time, \textcolor{blue}{a new method is proposed in this paper} to divide the individual search space according to the number of \textcolor{blue}{evolving} subpopulations, and also validating the previous approach.

%We will describe how we have tested our approach in the next Section.
\textcolor{blue}{The methodology to test our approach is described in the next Section.}

@@

\subsection{Proposed methods}
 \textcolor{blue}{Several implemented SLA algorithms are analyzed in this paper}
 by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, 
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. 
Therefore, a basic distributed NSGA-II algorithm without
overlapping sections \textcolor{blue}{has been used as the baseline option} (B).

>>>>

\newenvironment{bluecolor}{\par\color{black}}{\par}

@@

%%%NEW AND SHORTED
Parallel island-model co-evolutionary algorithms are well-known
methods, suitable for dealing with large multi-objective optimization
problems. \textcolor{black}{This paper proposes a version of these algorithms where} each island modifies a fragment of the
chromosome that encodes a possible solution to the
problem. \textcolor{black}{The objective of this paper is} to
demonstrate that automatically setting the size of the overlapping
fragments depending on the number of islands obtains better results
than using \textcolor{black}{a fixed overlapping} size. \textcolor{black}{This method has been
  compared} to other parallel evolutionary techniques considering a
different number of islands, chromosome sizes and benchmarks. The
analysis of the obtained experimental results, by using different
metrics, shows that our approach can provide statistically significant
improvements with respect to the base algorithm in high-dimensional,
un-decomposable, multi-objective problems. This opens a very promising
line to automatically adapt the overlapping sizes in this kind of
algorithms.

@@

% Maybe add a sentence explaining what problems are decomposable and which ones are not. - JJ
% Pablo: ok, done in blue
\begin{bluecolor}
One of the methods that are most extensively used  to deal with
high-dimensional, multi-objective optimization problems (MOPs) are
distributed co-evolutionary algorithms \cite{Gong15models}. However,
only decomposable problems \textcolor{black}{--that is, those whose  domain can be divided into subsets to calculate their partial solutions and combine them to form the solution of the original problem--,} could strictly be implemented in parallel
this way. A new parallel island-model evolutionary algorithm approach
\cite{Alba02Parallelism} is used here to deal with un-decomposable
problems: \textcolor{black}{instead of allocating disjoint subsets of components to the islands hosted by the available processors, it applies the variation operators only to the components  of the subset allocated to a given population, while keeping the rest of the components of the whole individual unaltered}. This avoids extra steps (and
subsequent communication overhead) to recompose the complete
individuals from other islands before evaluation. 
\end{bluecolor}

%1. MOP ARE
In MOPs, several \textcolor{black}{independent}
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives. The objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as possible. In order to do that\textcolor{black}{,} the
exploration of the search space must be very efficient, trying to
cover \textcolor{black}{all} possible exploration paths at the same time.

%2 COMPUTATION -> PARALLEL
This aspect of searching for non-dominated solutions, as well as, in
many cases, the size of the search space itself, implies a high demand
of computational time, leading to the proposal of parallel and distributed methods to solve them
\citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268},
%3 ONE WAY IS TO USE MOEAS BECAUSE
one of which \textcolor{black}{is} evolutionary algorithms (EAs)
\citep{DBLP:series/ncs/EibenS15}. EAs are bio-inspired meta-heuristics
that can be effectively used to find nearly optimal solutions for
optimization problems and are easily parallelizable, since they are
population-based. This means that creating a functionally equivalent parallel
version is as easy as distributing the population in communicating
computing nodes, which are usually called islands. The fact that
they are population-based helps find many solutions that compose the
Pareto front simultaneously.

Usually\textcolor{black}{,} an EA starts by generating \textcolor{black}{a \emph{population}, that is, a set of random solutions}, following a user-defined description. Then, it
evaluates each candidate solution, called \emph{individual}, assigning
it a \emph{fitness} \textcolor{black}{vector, in the case of MOPs} that describes how good the individual is
\textcolor{black}{solving} the target problem. New solutions are then generated by
the application of \emph{operators} that either mutate a single
solution or recombine different existing solutions. After each
iteration, called \emph{generation}, the least fit individuals are
removed, and the process continues until a user-defined stop condition
is met.

EAs \textcolor{black}{were} extended to multi-objective optimization through quite a number of Multi-Objective Evolutionary Algorithms (MOEAs) \cite{TalbiUnified2018}.
One of the best known and \textcolor{black}{highly} referenced MOEAs
\citep{Dorronsoro13superlinear} is the {\em 
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, which are divided into different ranks, based on 
dominance, to be selected for the next generations.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel,Gong15models}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies \textcolor{black}{using} different distribution
and sharing mechanisms, as there exists a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}. \textcolor{black}{This is also the case with the NSGA-II.} % Is that the case with NSGA? - JJ
% Please check the last question - JJ Pablo: done

Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}, 
meaning that the
MOEAs need to deal with large individuals \textcolor{black}{representing those variables} and spend significant extra
time \textcolor{black}{applying the} crossover, mutation and migration \textcolor{black}{operators}.
% If this is the only justification, it should have been assessed at the end. It would be better if it was qualified, in the sense that more exploration is needed or something like that - JJ
Different authors have 
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. % Why should it improve quality? - JJ Pablo: look at the 2 new paragraphs sent by Julio below (between \begin{bluecolor})
In this aspect, the
co-evolution model is a dimension-distributed model where a 
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. \textcolor{black}{To evolve the individuals of a subpopulation, the processor where it has been allocated can operate according to two main alternatives.} One
example of the application of the \textcolor{black}{first alternative} was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs \textcolor{black}{different alternatives for the recombination/cooperation of the subpopulations returned by worker processes}. A high
dimensional problem was used to compare these 
alternatives\textcolor{black}{; also, i}n \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although  both papers \textcolor{black}{described}
significant speedups, only a \textcolor{black}{small} number of nodes were
employed (8). However, when the number of islands is increased, the division
of \textcolor{black}{the portion of chromosome assigned to a given island} becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time. \textcolor{black}{Moreover, in this first alternative, each processor only works with the variables allocated to it of the individuals in its population. Thus, only whenever the cost function is decomposable in such a way that the variables allocated to different processors are not interdependent, the evaluation of the fitness of an individual in the population can be done by composing the cost evaluated separately by each processor from the values of the variables allocated to it. Otherwise, it has to be determined after composing the whole solution from the components allocated to each processor. In fact, different strategies to define the way to distribute the components into groups to be allocated to the available processors are required. They have been mainly defined according to three categories \cite{Jia19Coevo}: (1) static partitions of variables into groups; (2) random dynamic grouping that changes the variables among groups along the optimization process; and (3) learning-based dynamic grouping to distribute the components by applying techniques to detect the interdependence among variables along the process. }

@@

Another approach used to solve this kind of problems is the {\em
  SeLective operator Application} (SLA), presented in this paper, \textcolor{black}{can be considered as the second alternative to evolve the individuals of a subpopulation}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows \textcolor{black}{dealing} with un-decomposable problems.

@@

In our previous work \citep{Garcia16hpmoon} \textcolor{black}{this method was used} in a
preliminary experimental setup, although it is worth mentioning \textcolor{black}{that its name and acronym is established} in this paper for the first time.
\textcolor{black}{In that paper, it was shown}
that applying the variation operators only \textcolor{black}{to} specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, \textcolor{black}{the results showed} that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. 
This
motivate\textcolor{black}{d} us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more \textcolor{black}{extensive parameter evaluation}. %Again, "testing more stuff" is not a good motivation. You should try to address a specific research question that might have been raised in the previous paper - JJ
Besides comparing previous methods in this new experimental setup,  
a new method that automatically sets the size of the overlapped
sections depending on the number of available islands \textcolor{black}{is proposed in this paper, and compared} with previous versions.

@@

Results also show that the relation between the number of islands
(subpopulations) and the length of the chromosome (number of decision
variables) is also a relevant factor to determine the most efficient
alternative to \textcolor{black}{distribute} the decision variables: the higher the
dimensionality and the overlapping, the greater the benefit of
distribution in a large number of islands. This opens up a new
challenge to be addressed in the future, by experimenting with new
ways to calculate the overlapping sizes depending on different factors
and parameters.

The rest of the paper is organized as follows: after the State of the Art in distributed and co-evolutionary MOEAs, 
the \textcolor{black}{proposed} methodology and the compared algorithms are described in Section \ref{sec:methodology}. 
Then, the results of the experiments are presented (Section \ref{sec:results}), Finally, the conclusions and future work are discussed.

@@

%FERGU: Texto que me ha pasado Julio
%\textcolor{black}{Multi-objective optimization evolutionary algorithms (MOEAs) can take advantage of parallel and distributed processing \cite{Gong15models} not only to improve the quality of the solutions found, but also to speed up the execution times. Among the functional and data decomposition alternatives to parallelize an algorithm, parallel evolutionary algorithms usually implement data decomposition by identifying tasks that have to be repeatedly executed on different data. For example, in an evolutionary algorithm, the evaluation of the fitness function and the application of mutation operators to the individuals of the population can be independently applied to each individual. Moreover, in the context of data decomposition, two main alternatives can be distinguished. In the first one, the individuals of the population are distributed among several processes that independently compute their corresponding fitness functions. Once each process completes the fitness evaluation of the individuals allocated to it, another process applies the corresponding evolutionary operators to the individuals of the whole population to determine the population for the next generation. Then, the population is distributed again among the processes to obtain the new values of the fitness and so on until the end condition (a given number of generations, or a solution quality level) is met. It is clear that this alternative does not modify the convergence characteristics of the corresponding sequential evolutionary algorithm, although to complete each iteration communication is required between the process that applies the evolutionary operators to the population and the processes that evaluate the fitness values of the population. In the second alternative, several parallel processes execute iterations of the corresponding evolutionary algorithm on different subpopulations. In this case, the convergence behaviour could be different in sequential and parallel versions of the algorithms as the different subpopulations evolve in parallel and only exchange information about their individuals after completing a given number of generations. The number of individuals in the subpopulations (the grain size of the parallel procedure), and the characteristics of the communication (frequency of communication, character synchronous or asynchronous of the communication, characteristics of the communicated data, etc.) determine the performance of the parallel procedure. This second parallelization alternative (concurrent execution of evolutionary algorithms over multiple subpopulations) is the one considered in this paper. It can be implemented either by a master-worker approximation or by assigning each subpopulation to a different processor and communicating this processor with the others when it is required. In this second approach, according to the size of the grain (number of individuals in the subpopulations), two main alternatives can be also distinguished: the island parallel approach (coarse grain) and the cellular or diffusion approach (fine grain). In the master-worker approximation, the master distributes the subpopulations among the workers that communicate through the master. This way, as the communications increase (with the number of subpopulations and/or the frequency of communications among subpopulations) the master could become a bottleneck that reduces the efficiency of the parallel procedure.}

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other \textcolor{black}{kinds} of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, the authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population into elite and search sub-populations \citep{Wang09parallel}, or separating it \textcolor{black}{into} processors by objective \citep{Xiao03specialized}. For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector. Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Many real-world problems \textcolor{black}{require} high-dimensional data \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality. Other authors, such as Cao et al. \cite{CaoZLL17} separate the decision variables into several subpopulations after performing an interdependence analysis, but using a fixed number of processors.

The approach described by Dorronsoro et al. has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but implementing a
master-slave method that splits the population into several processors. As in previous works, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper \textcolor{black}{up to 128 islands have been used}.

Giving the responsibility of different parts of the search space to
different modules, has also been explored lately by some
researchers. For example, some authors focus on the solution space
\textcolor{black}{modelling} the PF by dividing it into different clusters, allowing an
easily parallelizable {\em divide and conquer} approach
\citep{cheng2015adaptive}. 
Even other types of metaheuristics, 
such as the Multi-Objective Ant Colony Optimization (MOACO) \cite{Mora13paretobased}
or Artificial Bee Colony Algorithm \cite{LUO2017235}, 
can rely on island models to divide the search space depending on 
the node to achieve better results.

In our previous work \citep{Garcia16hpmoon} some of
the aforementioned ideas \textcolor{black}{were used} to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections,  \textcolor{black}{with} $P$ being the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time, both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. \textcolor{black}{It was} discovered \textcolor{black}{in that paper} that the
performance using one or another method depends on \textcolor{black}{both} the number of
sections \textcolor{black}{individual's section used by each island} and the number of islands. This
motivated us to find a new automatic method to select this number of
sections of the chromosome to \textcolor{black}{be used}, depending on the number of
islands. Besides, previous experiments were performed on a single
processor \textcolor{black}{with a} synchronous island model and a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and \textcolor{black}{this time}, the experiments have been run on a parallel
cluster. 
Therefore, at the same time, \textcolor{black}{a new method is proposed in this paper} to divide the individual search space according to the number of \textcolor{black}{evolving} subpopulations, and also validating the previous approach.

%We will describe how we have tested our approach in the next Section.
\textcolor{black}{The methodology to test our approach is described in the next Section.}

@@

\subsection{Proposed methods}
 \textcolor{black}{Several implemented SLA algorithms are analyzed in this paper}
 by using a parallel multi-objective baseline algorithm
as a basis for comparisons. Our proposed methods, using different overlapping schemes, 
are based on NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. 
Therefore, a basic distributed NSGA-II algorithm without
overlapping sections \textcolor{black}{has been used as the baseline option} (B).
